URG analysis of the extended Kondo model with d-wave interaction

### 2025-12-10 submissions/manuscript_RP.tex
started creating referee response

\documentclass[%
 reprint,
superscriptaddress,
groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
aps,superscriptaddress,
pra,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-2}
% ,bm,color,hyperref
\usepackage{graphicx,braket,color,bm,amsmath,amssymb,hyperref}
% \graphicspath{{./figures/}}
\bibliographystyle{apsrev4-2}

\begin{document}

\title{Mott Criticality as the Confinement Transition of a Pseudogap-Mott Metal}

\author{Abhirup Mukherjee}
\email{am18ip014@iiserkol.ac.in}
\affiliation{Department of Physical Sciences, Indian Institute of Science Education and Research Kolkata, Nadia - 741246, India
}
\author{S. R. Hassan}
\email{shassan@imsc.res.in}
\affiliation{The Institute of Mathematical Sciences, HBNI, C.I.T. Campus, Chennai 600 113, India}
\author{Anamitra Mukherjee}
\email{anamitra@niser.ac.in}
\affiliation{School of Physical Sciences, National Institute of Science, Education and Research, HBNI, Jatni 752050, India}
\author{N. S. Vidhyadhiraja}
\email{raja@jncasr.ac.in}
\affiliation{Theoretical Sciences Unit, Jawaharlal Nehru Center for Advanced Scientific Research, Jakkur, Bengaluru 560064, India}
\author{A. Taraphder}
\email{arghya@phy.iitkgp.ac.in}
\affiliation{Department of Physics, Indian Institute of Technology Kharagpur, Kharagpur 721302, India}
\author{Siddhartha Lal}
\email{slal@iiserkol.ac.in}
\affiliation{Department of Physical Sciences, Indian Institute of Science Education and Research Kolkata, Nadia - 741246, India
}

\date{\today}

\begin{abstract}
The phenomenon of Mott insulation involves the localization of itinerant electrons due to strong local repulsion. Upon doping, a pseudogap (PG) phase emerges - marked by selective gapping of the Fermi surface without conventional symmetry breaking in spin or charge channels. A key challenge is understanding how quasiparticle breakdown in the Fermi liquid gives rise to this enigmatic state, and how it connects to both the Mott insulating and superconducting phases. Here, we develop a renormalization-based construction of strongly correlated lattice models that captures the emergence of the pseudogap phase and its transition to a Mott insulator. Applying a many-body tiling scheme to the fixed-point impurity model uncovers a lattice model with electron interactions and Kondo physics. At half-filling, the interplay between Kondo screening and bath charge fluctuations in the impurity model leads to Fermi liquid breakdown. This reveals a pseudogap phase characterized by a non-Fermi liquid (the Mott metal) residing on nodal arcs, gapped antinodal regions of the Fermi surface, and an anomalous scaling of the electronic scattering rate with frequency. The eventual confinement of holon–doublon excitations of this exotic metal obtains a continuous transition into the Mott insulator. Our results identify the pseudogap as a distinct long-range entangled quantum phase, and offer a new route to Mott criticality beyond the paradigm of local quantum criticality.

\end{abstract}

\maketitle
\section{Introduction}
The Landau quasiparticle paradigm~\cite{landau1959theory} underpins our understanding of metallic behaviour, where the screening of long-ranged repulsive interactions between delocalised electrons leads to well-defined excitations that carry the same quantum numbers as free electrons. This remarkable phenomenon - known as adiabatic continuity -  is captured by excitations described by poles of the single-particle Greens function of the interacting system. But what defines metallicity when quasiparticles are no longer well-defined~\cite{Phillips2013Unparticles}? Can a surface of zeros of the Greens function~\cite{dzyaloshinskii2003some}, rather than poles forming a Fermi surface, define an entirely new kind of metal? Could such an exotic state serve as the precursor to Mott localisation driven purely by repulsion, and without invoking magnetic order~\cite{Mott_1949}? These questions necessitate a radical departure in understanding conduction in correlated electron systems beyond the quasiparticle paradigm~\cite{phillips2022}. \\

\noindent
The anomalous pseudogap (PG) phase, observed in the cuprates, exhibits nodal-antinodal dichotomy with spectral weight concentrated on Fermi arcs around the nodal regions~\cite{loeserKapitulnik1996, Norman1998, Hashimoto2014}. In the PG phase of such Mott systems
~\cite{taillefer2010,FradkinRevModPhys2015,Sato2017,DoironLeyraud2017,Mukhopadhyay2019,Naman2021,Masafumi2025,KyungKotliar2006,MacridinAzevedo2006,sakai2009evolution,werner2009momentum,sakai2010doped,lin2010,gull2010momentum,gull2012,Mirzaei2013,gull2013,anirbanmott2,HilleAndergassen2020,Jiang2022,Scheurer2018,Krien2022,Kitatani2023,Sorella2023,Schafer2021,White1998,Ido2018,ProustTaillefer2019,Ponsioen2019,XuZhang2022}, the single-particle Green’s function develops surfaces of zeros, also known as Luttinger surfaces~\cite{Phillips2013,dzyaloshinskii2003some}, that signals a fundamental obstruction to quasiparticle formation. As interaction strength increases, these zero surfaces proliferate, fragmenting the Fermi surface into nodal arcs~\cite{sakai2009evolution,Phillips2013}, violating the Luttinger sum rule by reconfiguring Fermi volume~\cite{seki2017topological}, and breaking an emergent $\mathbb{Z}_2$ symmetry associated with the conservation of spin currents in Fermi liquids~\cite{Anderson2001,Huang2022}. Unlike conventional instabilities driven by poles, this phenomenon marks a topological breakdown of Fermi liquid theory~\cite{Coleman_2001,Phillips2013Unparticles}. We show that this breakdown arises from a strongly entangled, scale-invariant state of quantum matter, one that serves as the parent metal to the Mott insulator in a half-filled lattice model of strongly correlated electrons. This offers a concrete, controlled framework to understand both the non-quasiparticle character of the pseudogap phase and its continuous evolution into a repulsion-driven insulator, in full alignment with Mott's original vision~\cite{Mott_1949}.\\

\begin{figure*}[tbh]
    \centering
    \includegraphics[width=\linewidth]{Figure1_Nature.pdf}
    \caption{Schematic representation of (a) theoretical framework adopted by us (see main text for explanation of symbols), (b) phases obtained for the extended Hubbard model in two dimensions at $T=0$, and passage between them.}
    \label{schematic}
%    \label{schematic:b}
\end{figure*}

Our approach involves developing a renormalization group (RG)-based impurity-to-lattice framework that reconstructs the low-energy physics of a strongly correlated system from the fixed-point structure of a quantum impurity model (Fig.~\hyperref[schematic]{\ref*{schematic}(a)}). Starting from a dynamical Anderson impurity embedded in a correlated bath, we solve the system using a unitary RG (URG) approach~\cite{anirbanurg1,anirbanurg2}. We then construct a translation-invariant lattice Hamiltonian through a many-body tiling procedure that preserves non-local correlations and inherits momentum-resolved self-energies and renormalized couplings from the impurity solution. The resulting model is an \textit{extended}-Hubbard Hamiltonian on a two-dimensional square lattice at half-filling, with both on-site repulsion and nearest-neighbour spin exchange interactions. These correlations, derived directly from the impurity dynamics, enable tracking the evolution from the FL to the Mott insulator (MI) via an intermediate PG phase.

A central insight of our work is that the PG phase hosts a NFL state governed by Kondo frustration and doublon–holon deconfinement~\cite{keimer2015quantum} (Fig.~\hyperref[schematic]{\ref*{schematic}(b)}). A correlation scale ($W$) in the conduction bath suppresses Kondo screening (with coupling $J$) beyond a critical ratio of $W/J$, triggering the emergence of Luttinger surfaces in the lattice model. This breakdown of quasiparticles begins at the antinodes and progressively engulfs the nodal regions, driving a continuous transformation into the MI. The resulting NFL exhibits a pseudogapped density of states that vanishes as $\omega^{2}$ as $\omega \to 0$, multipartite long-range entanglement, and a universal scattering rate of the form $\sim (a + b\,\omega^2)^{-1}$ that grows as $\omega \to 0$. At the critical endpoint of this phase, the URG flow identifies a singular NFL described by the Hatsugai–Kohmoto (HK) model~\cite{Hatsugai1992,Baskaran1991,Huang2022}, with fully deconfined holon–doublon excitations.

The emergence of Luttinger surfaces alters the anomaly structure associated with the generalized symmetry of the FS~\cite{Altshuler_1998,Heath_2020,lanave2025}. The corresponding topological charge is defined via the Luttinger–Ward functional. This charge is modified by Green’s function zeros, signalling a shift in the underlying anomaly. Within our framework, this reorganization grants topological protection to the RG flows terminating in the PG regime, ensuring the stability of its NFL excitations. These are adiabatically connected to the Hatsugai–Kohmoto model~\cite{Hatsugai1992,Baskaran1991}, reinforcing the interpretation of the PG as a distinct quantum phase. We thus identify the NFL state as a new gapless, long-range entangled phase of correlated matter - the \textit{Mott metal} - whose deconfined holon-doublon excitations are eventually confined across the Mott transition~\cite{Mott_1949}.

While we demonstrate our approach for an extended Hubbard model, the underlying framework is broadly applicable to correlated quantum systems. The impurity-plus-tiling construction accommodates multi-orbital degrees of freedom, frustrated geometries, and nontrivial band topology. By deriving lattice Hamiltonians directly from renormalized local dynamics - without relying on mean-field approximations - this method preserves nonlocal correlations and enhances momentum-space resolution. It thus provides a versatile and controlled route for engineering strongly correlated models, particularly in regimes where PG formation, NFL scaling, and Mott physics intersect.

\section{Auxiliary Model and URG analysis}\label{auxurgsection}
To capture the interplay between Kondo physics and Mott localization, our approach relies on a two-dimensional auxiliary model consisting of an Anderson impurity embedded in a minimally-correlated bath, which we solve the system using a unitary RG (URG) approach~\cite{anirbanurg1,anirbanurg2}. The impurity model and URG analysis are laid out in this section. In order to make contact with a lattice model, we construct a translation-invariant Hamiltonian from the impurity model through a many-body tiling procedure: this preserves non-local correlations, as well as allows for momentum-resolved self-energies and renormalized couplings generated from the impurity solution. For the impurity model chosen in this work, tiling results in an \textit{extended}-Hubbard Hamiltonian on a two-dimensional square lattice at particle-hole symmetry (half-filling). The interaction terms within the lattice model are a result of analogous correlations present in the auxiliary model. This is laid out next in Section~\ref{tilingsection}.

\subsection{Lattice-embedded Impurity Model}
The auxiliary model consists of a correlated impurity site (with a double occupancy cost $U$) embedded within a correlated conduction bath defined on a square lattice (see Fig.~\hyperref[schematic]{\ref*{schematic}(a)}). The impurity site hybridises with the conduction bath sites adjacent to it through one-particle hybridisation $V$ as well as a spin-exchange interaction $J$ arising from fluctuations in local spin densities of the electrons. The conduction bath has minimal correlations in the form of a local correlation term $W$ on the sites that are adjacent to the impurity.

The Hamiltonian of the auxiliary model studied here is
\begin{equation}
\label{impurityModel}
\mathcal{H}_\text{aux} = H_\text{imp} + H_\text{coup} + H_\text{cbath},
\end{equation}
where the particle-hole symmetric impurity term is
\begin{equation}
H_\text{imp} = - \frac{U}{2} \left( \hat{n}_{d \uparrow} - \hat{n}_{d \downarrow} \right)^2,
\label{localrepulsion}
\end{equation}
describing on-site Coulomb interactions at the impurity. Here, $c^\dagger_{d\sigma}$ creates an electron with spin $\sigma$ at the impurity site, and $\hat{n}_{d\sigma} = c^\dagger_{d\sigma} c_{d\sigma}$ denotes the corresponding number operator.

The impurity couples to four nearest-neighbour bath sites $c_{Z\sigma}$ through both hybridization and Kondo exchange:
\begin{equation}\begin{aligned}
H_\text{coup} = \frac{J}{2} \hspace{-0.25cm}\sum_{\sigma_1,\sigma_2,Z} 
%\sum_{Z} 
{\bf S}_d \cdot c^\dagger_{Z\sigma_1} {\boldsymbol \tau}_{\sigma_1 \sigma_2} c_{Z\sigma_2} 
- V \sum_{\sigma,Z} \left( c^\dagger_{Z\sigma} c_{d\sigma} + \text{h.c.} \right),    
\end{aligned}
\label{impbathcoup}\end{equation}
where ${\boldsymbol \tau}$ are the Pauli matrices, and $Z$ indexes the four neighbouring bath sites. The half-filled bath includes nearest neighbour hopping and on-site correlations on the site neighbouring the impurity:
\begin{equation}
H_\text{cbath} = \sum_{\bf k} \epsilon_{\bf k} c^\dagger_{{\bf k},\sigma} c_{{\bf k},\sigma} - \frac{W}{2} \sum_{Z} \left( \hat{n}_{Z\uparrow} - \hat{n}_{Z\downarrow} \right)^2,
\label{bathke}
\end{equation}
where $\epsilon_{\bf k} = -2t(\cos k_x + \cos k_y)$, and $W$ parametrizes local repulsion in the bath that frustrates Kondo screening and drives the system towards local-moment formation. A crucial feature of this construction is that the Kondo coupling acquires a momentum structure upon Fourier transforming:
\begin{equation}
J_{{\bf k}, {\bf k}^\prime} = \frac{J}{2} \left[ \cos(k_x - k^\prime_x) + \cos(k_y - k^\prime_y) \right],
\end{equation}
which respects the $C_4$ symmetry of the square lattice.

\subsection{The unitary renormalisation group method}
In order to obtain the various low-energy phases of our impurity model, we perform a scaling analysis of the associated Hamiltonian using the recently developed unitary renormalisation group (URG) method ~[19,39,40].
%\cite{anirbanurg1,anirbanurg2}. 
The method has been applied successfully on a wide variety of problems of correlated fermions~\cite{santanukagome,1dhubjhep1,anirbanmott1,siddharthacpi,anirban_kondo,Patra_2023}. The method proceeds by resolving quantum fluctuations in high-energy degrees of freedom, leading to a low-energy Hamiltonian with renormalised couplings and new emergent degrees of freedom. Typically, for a system with Fermi energy \(\epsilon_F\) and bandwidth \(E_N\), the sequence of isoenergetic shells \(\left\{E_{(j)}\right\}, E_{j}\in \left[E_0, E_N\right] \) define the states whose quantum fluctuations we sequentially resolve. The momentum states lying on shells \(E_N\) that are far away from the Fermi surface comprise the UV states, while those on shells near the Fermi surface comprise the IR states.

As a result of the URG transformations, the Hamiltonian \(H_{(j)}\) at a given RG step \(j\) involves scattering processes between the \(k-\)states that have energies lower than \(D_{(j+1)}\). The unitary transformation \(U_{(j)}\) is then defined so as to remove the number fluctuations of the currently most energetic set of states \(D_{(j)}\)~\cite{anirbanurg1,anirbanurg2}:
\begin{eqnarray}
	H_{(j-1)} = U_{(j)} H_{(j)} U^\dagger_{(j)}~, \text{such that} ~\left[H_{(j-1)}, \hat n_{j}\right] =0~.
\end{eqnarray}
The eigenvalue of $\hat{n}_{j}$ has, thus, been rendered an integral of motion (IOM) under the RG transformation.

The unitary transformations can be expressed in terms of a generator \(\eta_{(j)}\) that has fermionic algebra~\cite{anirbanurg1,anirbanurg2}:
\begin{eqnarray}
	\label{unitary}
	U_{(j)} = \frac{1}{\sqrt 2}\left(1 + \eta_{(j)} - \eta_{(j)}^\dagger\right)~,~ \quad\left\{ \eta_{(j)},\eta_{(j)}^\dagger \right\} = 1~,
\end{eqnarray}
where \(\left\{\cdot\right\}\) is the anticommutator. The unitary operator \(U_{(j)}\) that appears in Eq.~\eqref{unitary} can be cast into the well-known general form \(U = e^\mathcal{S}, \mathcal{S} = \frac{\pi}{4}\left( \eta^\dagger_{(j)} - \eta_{(j)} \right)\) that a unitary operator can take, defined by an anti-Hermitian operator \(\mathcal{S}\). The generator \(\eta_{(j)}\) is given by the expression\cite{anirbanurg1,anirbanurg2}
\begin{eqnarray}
	\eta^\dagger_{(j)} = \frac{1}{\hat \omega_{(j)} - \text{Tr}\left(H_{(j)} \hat n_{j}\right) } c^\dagger_{j} \text{Tr}\left(H_{(j)}c_{j}\right)~.
\end{eqnarray}
The operators \(\eta_{(j)},\eta^\dagger_{(j)}\) behave as the many-particle analogues of the single-particle field operators \(c_j,c^\dagger_j\) - they change the occupation number of the single-particle Fock space \(\ket{n_j}\).  The important operator \(\hat \omega_{(j)}\) originates from the quantum fluctuations that exist in the problem because of the non-commutation of the kinetic energy terms and the interaction terms in the Hamiltonian:
\begin{eqnarray}
	\hat \omega_{(j)} = H_{(j-1)} - H^i_{(j)}~.
	\label{omega}
\end{eqnarray}
\(H^i_{(j)}\) is the part of \(H_{(j)}\) that commutes with \(\hat n_j\) but does {\it not} commute with at least one \(\hat n_l\) for \(l < j\). The RG flow continues up to energy \(D^*\), where a fixed point is reached from the vanishing of the RG function. 
Detailed comparisons of the URG with other methods (e.g., the functional RG, spectrum bifurcation RG etc.) can be found in Refs.~\cite{anirbanmott1,anirbanurg1}, and implementation on the present problem in  Appendix A.

\section{Tiling formalism}\label{tilingsection}
\subsection{Reconstructing of lattice model from auxiliary model}
To reconstruct a translationally invariant lattice Hamiltonian from the impurity model (with Hamiltonian $\mathcal{H}_\text{aux}$), we define a many-body tiling prescription that embeds the auxiliary impurity system uniformly across the two-dimensional square lattice. Thus, the tiling process systematically derives the full lattice dynamics from the renormalized fixed-point structure of the impurity. As described in Section~\ref{auxurgsection}, the auxiliary model consists of a single impurity embedded in a correlated bath at the location $\mathbf{r}_d$: 
\begin{equation}
H(\mathbf{r}_d) = H_{\text{imp}} + H_{\text{cbath}} + H_{\text{imp-cbath}}~.
\label{auxmodel}
\end{equation}
The form of various terms in eq.\eqref{auxmodel}
have been defined in eqs.\eqref{localrepulsion}, \eqref{impbathcoup} and \eqref{bathke} in Section~\ref{auxurgsection}.
% where the impurity, conduction bath and impurity–bat Hamiltonians are defined as
% \begin{equation}
% \begin{aligned}
% H_{\text{imp}} &= -\frac{U}{2} \left( \hat{n}_{\mathbf{r}_d\uparrow} - \hat{n}_{\mathbf{r}_d\downarrow} \right)^2,\\
% H_{\text{cbath}} &= -\frac{1}{\sqrt{\mathcal{Z}}} t \sum_{\langle \mathbf{r}_i, \mathbf{r}_j \rangle \ne \mathbf{r}_d; \sigma} \left( c^\dagger_{\mathbf{r}_i\sigma} c_{\mathbf{r}_j\sigma} + \text{h.c.} \right) \\
% &- \frac{1}{2\mathcal{Z}} W \sum_{\mathbf{z} \in \text{NN}(\mathbf{r}_d)} \left( \hat{n}_{\mathbf{z}\uparrow} - \hat{n}_{\mathbf{z}\downarrow} \right)^2,\\
% H_{\text{imp-cbath}} &= \frac{J}{\mathcal{Z}} \sum_{\sigma, \sigma'} \sum_{\mathbf{z} \in \text{NN}(\mathbf{r}_d)} \mathbf{S}_{\mathbf{r}_d} \cdot \boldsymbol{\tau}_{\sigma \sigma'} c^\dagger_{\mathbf{z}\sigma} c_{\mathbf{z}\sigma'} \\
% &- \frac{V}{\sqrt{\mathcal{Z}}} \sum_{\sigma} \sum_{\mathbf{z} \in \text{NN}(\mathbf{r}_d)} \left( c^\dagger_{\mathbf{r}_d\sigma} c_{\mathbf{z}\sigma} + \text{h.c.} \right).
% \end{aligned}
% \end{equation}
% $\boldsymbol{\tau} = (\tau_x, \tau_y, \tau_z)$ are Pauli matrices, and ${\bf r}_d$ denotes the impurity position.

The tiled lattice Hamiltonian is generated by symmetrically translating this impurity model:
\begin{equation}
\mathcal{H}_{\text{tiled}} = \mathcal{T}[H(\mathbf{r}_d)] = \sum_{\mathbf{r}} T^\dagger(\mathbf{r} - \mathbf{r}_d) H(\mathbf{r}_d) T(\mathbf{r} - \mathbf{r}_d),
\end{equation}
where $T(\mathbf{a})$ are many-body translation operators~\cite{stoyanova}:
\begin{equation}
T^\dagger(\mathbf{a}) \mathcal{O}(\{\mathbf{r}_i\}) T(\mathbf{a}) = \mathcal{O}(\{\mathbf{r}_i + \mathbf{a}\}).
\end{equation}
Applying this to each term yields:
\begin{equation}\hspace{-0.5cm}\begin{aligned}
\mathcal{T}[H_{\text{cbath}}] &= -\frac{(N - 2)t}{\sqrt{\mathcal{Z}}} \sum_{\langle \mathbf{r}_i, \mathbf{r}_j \rangle; \sigma} \left( c^\dagger_{\mathbf{r}_i\sigma} c_{\mathbf{r}_j\sigma} + \text{h.c.} \right) \\
&- \frac{1}{2} W \sum_{\mathbf{r}} \left( \hat{n}_{\mathbf{r}\uparrow} - \hat{n}_{\mathbf{r}\downarrow} \right)^2, \\
\mathcal{T}[H_{\text{imp}}] &= -\frac{U}{2} \sum_{\mathbf{r}} \left( \hat{n}_{\mathbf{r}\uparrow} - \hat{n}_{\mathbf{r}\downarrow} \right)^2, \\
\mathcal{T}[H_{\text{imp-cbath}}] &= \sum_{\langle \mathbf{r}_i, \mathbf{r}_j \rangle} \left[ \frac{2J}{\mathcal{Z}} \mathbf{S}_{\mathbf{r}_i} \cdot \mathbf{S}_{\mathbf{r}_j} - \frac{2V}{\sqrt{\mathcal{Z}}} \sum_\sigma \left( c^\dagger_{\mathbf{r}_i\sigma} c_{\mathbf{r}_j\sigma} + \text{h.c.} \right) \right],
\end{aligned}\end{equation}
with local spin operators defined as
\begin{equation}
\mathbf{S}_{\mathbf{r}} = \sum_{\sigma, \sigma'} c^\dagger_{\mathbf{r}\sigma} \boldsymbol{\tau}_{\sigma \sigma'} c_{\mathbf{r}\sigma'}.
\end{equation}

To avoid overcounting of the non-interacting bath terms, we subtract extra copies of the kinetic energy term. The reconstructed extended–Hubbard model then reads:
\begin{equation}
\begin{aligned}
\mathcal{H}_{\text{tiled}} = & -\frac{1}{\sqrt{\mathcal{Z}}} \tilde{t} \sum_{\langle \mathbf{r}_i, \mathbf{r}_j \rangle; \sigma} \left( c^\dagger_{\mathbf{r}_i\sigma} c_{\mathbf{r}_j\sigma} + \text{h.c.} \right) \\
&+ \frac{\tilde{J}}{\mathcal{Z}} \sum_{\langle \mathbf{r}_i, \mathbf{r}_j \rangle} \mathbf{S}_{\mathbf{r}_i} \cdot \mathbf{S}_{\mathbf{r}_j} - \frac{1}{2} \tilde{U} \sum_{\mathbf{r}} \left( \hat{n}_{\mathbf{r}\uparrow} - \hat{n}_{\mathbf{r}\downarrow} \right)^2,
\end{aligned}
\end{equation}
where the effective lattice couplings are given by:
\begin{equation}
\tilde{t} = t + 2V,\quad \tilde{U} = U + W,\quad \tilde{J} = 2J.
\end{equation}
% \end{widetext}

A non-zero chemical potential in the bath and/or the impurity can be included to study the effects of doping. Crucially, the eigenstates of $\mathcal{H}_\text{tiled}$ obey a many-body generalization of Bloch’s theorem~\cite{stoyanova}, which allows for the exact computation of momentum-resolved observables (presented here for a $77\times 77$ {\bf k}-space Brillouin zone grid). We shall demonstrate below that the lattice embedding and tiling procedures provide a controlled route to capture low-energy NFL behavior and Mott criticality directly from a quantum impurity model.

\subsection{Mapping eigenstates using a manybody Bloch's theorem}
As mentioned above, the eigenstates \(\ket{\Psi_{s}}\) (\(s = \left({\bf k}, n\right)\)) of the lattice Hamiltonians obtained using  the tiling procedure enjoy a {\it many-body} Bloch's theorem~\cite{stoyanova}, as the tiling procedure restores the translation symmetry of the Hamiltonian. This means that the {\it local} eigenstates \(\ket{\psi_n\left({\bf r}_d\right)}\) (with the impurity located at an arbitrary position \({\bf r}_d\)) of the unit cell auxiliary model Hamiltonian \(\mathcal{H}_\text{aux}({\bf r}_d)\) can be used to construct eigenstates of the lattice Hamiltonian. The index \(n(=0,1,\ldots)\) in the subscript indicates that it is the \(n^\text{th}\) eigenstate of the auxiliary model.

The state \(\ket{\psi_n\left({\bf r}_d\right)}\) does not specify the position of the zeroth site, because the unit cell Hamiltonian \(\mathcal{H}_\text{aux}({\bf r}_d)\) itself has been averaged over \(\mathcal{Z}\) zeroth sites. Accordingly, we can express the averaged eigenstate \(\ket{\psi_n\left({\bf r}_d\right)}\) as
\begin{equation}\begin{aligned}
	\ket{\psi_n\left({\bf r}_d\right)} = \frac{1}{\sqrt\mathcal{Z}}\sum_{{\bf z} \in \text{NN}({\bf r}_d)}\ket{\psi_n\left({\bf r}_d, {\bf z}\right)}~,
\end{aligned}\end{equation}
where \(\ket{\psi_n\left({\bf r}_d, {\bf z}\right)}\) is an auxiliary model eigenstate with the impurity and zeroth sites placed at \({\bf r}_d\) and \({\bf z}\). With this in mind, the following unnormalised combination of the auxiliary model eigenstates satisfies a many-particle equivalent of Bloch's theorem~[53]:
%\cite{stoyanova}:
\begin{equation}\begin{aligned}\label{eigenstateProposal1}
	\ket{\Psi_{s}} \equiv \ket{\Psi_{{\bf k}, n}} &= \frac{1}{\sqrt N}\sum_{{\bf r}_d} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_{n}\left({\bf r}_d\right)} \\
\end{aligned}\end{equation}
where \(N\) is the total number of lattice sites and \({\bf r}_d\) is summed over all lattice spacings. The set of \(n=0\) states form the lowest band in the spectrum of the lattice, while higher values of \(n\) produce the more energetic bands. The ground state \(s = s_0\) is obtained by setting \({\bf k}\) and \(n\) to 0:
\begin{equation}\begin{aligned}\label{groundstateProposal1}
	\ket{\Psi_\text{gs}} = \frac{1}{\sqrt N}\sum_{{\bf r}_d} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_\text{gs}\left({\bf r}_d\right)}
\end{aligned}\end{equation}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig2.pdf}
    \caption{(a) Phase diagram of impurity model at strong coupling in $U$ in terms of competing dimensionless Kondo ($J/t$) and bath correlation ($W/t$) couplings. Colourbar represents the fraction of states on the Fermi surface replaced by zeros of the Greens function (Luttinger surface). A pseudogap phase (shaded region) is observed between the local Fermi liquid (red region) and local moment (blue region) phases (b) $k$-space-resolved spin-spin correlation $\chi_s(d,{\bf k}) = \braket{{\bf S}_d\cdot{\bf S}_{\bf k}}$ in the pseudogap phase of the impurity model. Antinodal regions are observed to decouple from Kondo screening of the impurity. (c) Upon tiling, this leads to a $k$-space-resolved antinodal gap in the electronic density of states of the lattice model, corresponding to Luttinger surfaces of zeros.}
    \label{spinCorr}
%    \label{spinCorr:b}
%    \label{spinCorr:c}
\end{figure*}

\subsection{Mapping the retarded Greens function from auxiliary model to lattice model}
We now show how to relate the one-particle retarded Greens functions of the bulk lattice to that of the auxiliary model. The retarded time-domain lattice \(k-\)space Greens function is defined as
\begin{equation}\begin{aligned}
	\tilde G({\bf K}\sigma; t) = -i\theta(t) \braket{\Psi_\text{gs} | \left\{ c_{{\bf K}\sigma}(t), c^\dagger_{{\bf K}\sigma} \right\} | \Psi_\text{gs}}~.
\end{aligned}\end{equation}
where the bulk Hamiltonian \(H_\text{tiled}\) leads to the dynamics of the annihilation operators at time \(t\): 
\begin{equation}\begin{aligned}\label{heisenberg}
	c_{{\bf K}\sigma}(t) = e^{it H_\text{tiled} }c_{{\bf K}\sigma}e^{-i t H_\text{tiled}}~.
\end{aligned}\end{equation}
We now proceed to simplify one of the terms of the anticommutator (for simplicity of notation):
\begin{equation}\begin{aligned}\label{greensFunction1}
	&\braket{\Psi_\text{gs} | c_{{\bf K}\sigma}(t) c^\dagger_{{\bf K}\sigma} | \Psi_\text{gs}} \\
    &= \frac{1}{N^2}\sum_{\vec r,\vec \Delta}e^{-i{\bf K}_0\cdot\vec\Delta}\braket{\psi_{0}(\vec r+\vec \Delta) | c_{{\bf K}\sigma}(t) c^\dagger_{{\bf K}\sigma} | \psi_{0}(\vec r)}~.
\end{aligned}\end{equation}
Using eq.~\eqref{eigenstateProposal1}, eq.~\eqref{greensFunction1} becomes
\begin{widetext}
	\begin{equation}\begin{aligned}\label{greensfunction2}
		\braket{\Psi_\text{gs} | c_{{\bf K}\sigma}(t) c^\dagger_{{\bf K}\sigma} | \Psi_\text{gs}} = \frac{1}{N^2}\sum_s\sum_{\vec r,\vec \Delta}\sum_{\vec r^\prime,\vec \Delta^\prime}e^{-i\vec k_0\cdot\vec\Delta} e^{i\vec k\cdot\vec\Delta^\prime}\braket{\psi_{0}(\vec r+\vec \Delta) | c_{{\bf K}\sigma}(t) \ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)} c^\dagger_{{\bf K}\sigma} | \psi_{0}(\vec r)}~.
\end{aligned}\end{equation}
In order to bring this expression closer to the form of an auxiliary model Greens function, we
\begin{itemize}
    \item use the relation $\ket{\psi(\vec r + \Delta)} = T^\dagger(\vec \Delta)\ket{\psi(\vec r)}$, where \(T^\dagger(\vec \Delta)\) translates all lattice sites by the vector \(\vec\Delta\)~,
    \item use the property $T\left(\vec a\right) c({\bf K}) T^\dagger\left(\vec a\right) = e^{-i{\bf K}\cdot\vec a}c({\bf K})$~, and
    \item make the substitution \(\vec \Delta^\prime \to \vec \Delta^\prime + \vec \Delta\)~.
\end{itemize}
This leads to the expression
\begin{equation}\begin{aligned}\label{greensfunction3}
		\braket{\Psi_\text{gs} | c_{{\bf K}\sigma}(t) c^\dagger_{{\bf K}\sigma} | \Psi_\text{gs}} = \frac{1}{N}\sum_n\sum_{\vec r,\vec r^\prime,\vec \Delta^\prime} e^{i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\braket{\psi_{0}(\vec r) | c_{{\bf K}\sigma}(t) \ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)} c^\dagger_{{\bf K}\sigma} | \psi_{0}(\vec r)}~,
\end{aligned}\end{equation}
where the sum over \(s=(\vec k, n)\) has been reduced to a sum over the auxiliary model eigenstate index \(n\) because of the Kronecker delta \(\delta\left(\vec k_0 + {\bf K} - \vec k\right)\). This can be further simplified by splitting the sum over \(\vec\Delta^\prime\) into positive and negative parts and then making the transformation \(\vec r^\prime \to \vec r^\prime + \vec \Delta^\prime\):
\begin{equation}\begin{aligned}
	&\sum_{\vec r^\prime,\vec \Delta^\prime} e^{i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)}= \frac{1}{2}\sum_{\vec r^\prime,\vec \Delta^\prime}\left[ e^{i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)} + e^{-i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\ket{\psi_n(\vec r^\prime-\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)}\right]~, \\
																																												   &= \frac{1}{2}\sum_{\vec r^\prime,\vec \Delta^\prime}\left[ e^{i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)} + e^{-i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\ket{\psi_n(\vec r^\prime)}\bra{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\right]~. \\
\end{aligned}\end{equation}
For each pair of \(\vec r^\prime\) and \(\vec\Delta^\prime\), the term within the box brackets has the form of a two-level Hamiltonian between the states \(\ket{\psi_n(\vec r^\prime)}\) and \(\ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\), with a tunnelling amplitude \(e^{i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\). The term can therefore be written in the eigenbasis of this Hamiltonian:
\begin{equation}\begin{aligned}

>>>>

Precisely at $W/J=(W/J)_{\text{PG}}$, the antinode joins this connected region of zeros in $J_{{\bf k}_1, {\bf k}_2}$, marking the decoupling of the antinodes from all other points in the neighbourhood of the FS. This is an interaction-driven Lifshitz transition of the FS, and marks the entry into a PG phase possessing Fermi arcs~\cite{WuFerrero2018}. Importantly, it coincides with an emergent two-channel Kondo (2CK) impurity model, where each channel corresponds to a pair of Fermi arcs on opposite faces of the conduction bath FS. The 2CK nature of the PG is guaranteed by the symmetry of $J_{{\bf k},{\bf k'}}$:
